
---
title: "Git & SQL Practical Assignment"
author: "Student Name"
output:
  pdf_document: default
  html_document: default
---

## Assignment Overview

This assignment guides you through using Git and SQL in RStudio. You will clone a repository, work with a SQLite database, and use RStudio's Git interface to manage your work.

---

## Step 1: Clone the Repository Using RStudio

1. Open RStudio.
2. Go to **File** → **New Project** → **Version Control** → **Git**.
3. Paste the repository URL provided by your instructor.
4. Choose a local folder and click **Create Project**.

---

## Step 2: Copy the R Markdown Template

1. In the **Files** pane, locate `assignment_template.Rmd`.
2. Right-click and select **Copy**.
3. Rename the copy to `assignment_<your_student_id>.Rmd` (replace `<your_student_id>` with your actual ID).

---

## Step 3: Open Your Rmd File

Double-click your new `.Rmd` file to open it in the editor.

---

## Step 4: Connect to the SQLite Database

```{r setup, message=FALSE}
library(DBI)
library(RSQLite)
con <- dbConnect(RSQLite::SQLite(), "output_database.sqlite")
```

---

## Step 5: Perform SQL Queries
a. List all tables in the database

```{r}

dbListTables(con)

```

b. Show the first 5 rows o a table

```{r}

dbGetQuery(con, "SELECT * FROM mortality_clean LIMIT 5;")

```

c. Count the number of records in another table 

```{r}

dbGetQuery(con, "SELECT COUNT(*) FROM demographics_clean;")

```
d. Join two tables and display the first 5 results

```{r}


# dbGetQuery(con, "
#   SELECT Patients.PatientID, Name, Diagnosis
#   FROM Patients
#   JOIN Diagnoses ON Patients.PatientID = Diagnoses.PatientID
#   LIMIT 5;
# ")


```


---

## Step 6: Disconnect from the Database
```{r}

dbDisconnect(con)

```

---

## Step 7: Commit and Push Your work Using RStudio
1. In the Git pane, check the box next to your .Rmd file.
2. Click Commit.
3. Enter a commit message (e.g., “Completed assignment by <your_student_id>”).
4. Click Commit.
5. Click Push to upload your changes to the remote repository.



